Supporting Information for\\ ``A weakly structured stem for human origins in Africa''

### 2022-07-27 supplement.tex
clean up fixed parameters section

In our preliminary models, the unconstrained inferred split times between East
African populations and Europeans was inferred to have occurred around $40$ka,
which is more recent than we may have expected given the well-documented
presence of modern humans outside of Africa at that time \citep{}, as well as
previous genetics-based dating of splits between African and Eurasian
populations \citep{}. The unconstrained inferred split between East and West
African populations occurred about $10$ka earlier
(Section~\ref{sec:supp-results}).
%We see two non-exclusive explanations for this apparent discrepancy. First,
%there is uncertainty in our model fit of this parameter. The $95\%$ confidence
%interval allows for a split as far back as \sgcomment{$XX$}.
Across all models, the $95\%$ confidence intervals for the OOA event was
roughly $38 \pm 5$ka, and for the East/West African split was $45 \pm 10$ka.
However, our confidence intervals do not capture all sources of uncertainty.
For example, model mis-specification can lead to a mismatch between an inferred
split time and the first time populations actually split. Our model supposes a
single OOA event followed by constant symmetric migrations between populations,
which is certainly an oversimplification \citep{}. Additionally, most previous
estimates of the time of the OOA expansion compared Eurasian and West African
populations. Our inferred time of the East/West African divergence aligns with
previous estimates of the OOA event at $\approx 60$ka (themselves inferred from
oversimplified demographic models). Rather than adding the many parameters
required to account for these effects, we fixed the split time between
Europeans and the East African ancestral population at $50$ka, and the split
between East/West African population at $60$ka.
%We discuss the sensitivity of our model to these assumptions in Section
%\ref{sec:change-dates}.

@@

\subsubsection{\aprcomment{TODO:} High-coverage Nama samples}

@@

All supplementary inference results are provided in Demes format
\citep{Gower2022-yn} as a supplemental zipped archive.

>>>>

In our preliminary models, the unconstrained inferred split times between East
African populations and Europeans was inferred to have occurred by around
$40$ka, which is more recent than we may have expected given the
well-documented presence of modern humans outside of Africa at that time
\citep[e.g.]{Hublin2020-nk,Hajdinjak2021-jo}, as well as previous
genetics-based dating of splits between African and Eurasian populations
\citep[e.g.]{Jouganous2017-pq,Kamm2020-vn}. The unconstrained inferred split
between East and West African populations occurred about $10$ka earlier. These
recent split dates seems to be at odds with archaeological evidence for earlier
Out-of-Africa (OOA) expansion, and with the estimated date for Neanderthal
admixture into OOA populations. The parameter uncertainty due to finite genomes
does not account for this difference: in the high-likelihood models, the $95\%$
confidence intervals for the East African-European split was roughly $38 \pm
5$ka, and for the East/West African split was $45 \pm 10$ka, with the
uncertainties reflecting 2 standard deviations according to bootstrap of
genomic regions.  Thus we expect these discrepancies to result from model
misspecification in the details of the divergence between the ancestors of OOA
and east African populations. Our models supposes a single split prior to the
OOA event followed by constant symmetric migrations between eastern and OOA
populations, which is certainly an oversimplification.

Since a detailed model for the Out of Africa and Back-to-Africa was not our
main focus, we chose to keep this simplified model, but fix the split between
east African-European split at $50$ka. This allowed us to include Neanderthal
admixture into the OOA population at the previously-estimated $45$ka. Fixing
this parameter pushed back the estimated value for the split between the
ancestors of East and West African populations at around $60$ka, a value that
we took as fixed in most subsequent analyses. 
We discuss the sensitivity of our model to these assumptions in Section
\ref{sec:change-dates}.

%\sgcomment{I feel like we should improve the discussion of relationship
%with previous estimates below, or remove.}
%
%In addition to archaeological evidence, previous genetic dating suggested 
%earlier dates for a split between European and African populations. Most of 
%these results focused on populations of recent West African ancestry. 
%
%Additionally, most previous
%estimates of the time of the OOA expansion compared Eurasian and West African
%populations. Our inferred time of the East/West African divergence aligns with
%previous estimates of the OOA event at $\approx 60$ka (themselves inferred from
%oversimplified demographic models).

@@

@@

All supplementary inference results are provided in Demes format
\citep{Gower2022-yn} as a supplemental zipped archive. For each scenario,
we refit the continuous migration and multiple merger models, with
stem migration.
